### 2004-09-29 classes/hopl/control/control.tex

\documentclass[12pt]{article}

\usepackage{amssymb}
\usepackage[all]{xy}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{enumerate}

\title{Modern Control Operators}
\author{Sam Tobin-Hochstadt}
\date{\today}

\input xy
\xyoption{all}

\begin{document}
\bibliographystyle{plain}

\maketitle

\makebox{\hspace{21mm}
\xymatrix{{}& *++[F--]{~~~~~\cite{monads}~~~~~}& *++[F--]{~~~~~\cite{sitaram94thesis}~~~~~} & *++[F--]{~~~~~\cite{luc94partial}~~~~~}\\
& {} & *++[F]{~~~~~\cite{sitaram93handling}~~~~~}\ar[u] & *++[F--]{~~~~~\cite{queinnec93library}~~~~~} \ar[u]
\\ &{} & {} &*++[F]{~~~~~\cite{queinnec91dynamic}~~~~~} \ar[u] 
\\ &{} & {}& *++[F]{~~~~~\cite{hieb90continuations}~~~~~} \ar@/^/[uul]  \ar[u]   \\
& *++[F]{~~~~~\cite{danvy90abstracting}~~~~~} \ar@/^/[urr]  \ar[uuuu] \ar@//[uurr] \ar@//[uuurr] &*++[F]{~~~~~\cite{sitaram90control}~~~~~} \ar[ur] \ar[uuu] \ar[l] \ar@/^/[uuur]\\&
  {}&*++[F]{~~~~~\cite{felleisen87syntactic}~~~~~} \ar[u] \\
  &{}&*++[F--]{~~~~~\cite{felleisen86secd}~~~~~} \ar[u] \\
}
}

\subsection*{A Syntactic Theory of Sequential Control} 
Eugene Kohlbecker, Daniel Friedman, Matthias Felleisen and Bruce Duba 1987
\cite{felleisen87syntactic}

This paper lays out a semantics for higher order control based on
syntactic transformation of evaluation contexts.  Semantic analysis
with syntactic transformation was introduced primarily by Plotkin
\cite{gdp:cbn-cbv}, and evaluation contexts by Felleisen in
\cite{felleisen86secd}.  They present the control operators ${\mathcal
A}$ and ${\mathcal C}$, corresponding to {\tt abort} and {\tt
control}.  They then provide a context based semantics for these
operators, and prove a number of important properties, such as
Church-Rosser, for the new extended language, called $\lambda_C$.

This paper, referenced by many of the following papers in the area, lays
the groundwork for future work in the study of higher order control
operators.  It summarizes work published in a number of other settings
by the authors, places the previous results in a self-contained
context, and proves a number of key theorems about the system.

\subsection*{Control Delimiters and their Hierarchies}
Matthias Felleisen and Dorai Sitaram 1990 \cite{sitaram90control}

In this paper, Felleisen and Sitaram describe a new pair of control
operators, providing delimited continuations, first described in \cite{mf-prompts}.
Such continuations do not capture the entire rest of the computation,
but merely that back to an enclosing prompt.  Further, instead of
using the standard Scheme control operator {\tt call/cc}, they use the
operator $\mathcal{F}$, also previously defined by Felleisen.  This
operator is simpler, but provides all the power of the others.

The authors then provide a number of potential applications for such a
control operator, as well as implementation strategies.  Some of the
potential applications include coroutines, engines and implementations
of first-order control handlers such as exceptions.  Finally, the
authors discuss the possibilites of generating hierarchies of control
delimiters, so that continuations can be captured back to specific
delimiters, rather than the nearest enclosing delimiter.  They show
how this is easily expressible in terms of the original operators.

The sequence of papers culminating in this one introduced the delimited
continuation concept as a research topic.  Sitaram, Danvy and other
built on this work to create other, potentially more sophisticated
control operators.  However, the early papers by Felleisen and
collaborators represented their introduction.

\subsection*{Abstracting Control}
Olivier Danvy and Andre Filinski 1990 \cite{danvy90abstracting}

This paper is at once both very similar to, and very different from,
the previous paper.  It is similar in that it too introduces a new
control operator for delimited continutation, here called {\tt shift}
and {\tt reset}.  Further, shift and reset are very similar to the
previous control operators introduced by Sitaram and Felleisen.
Similar enough, in fact, that it takes serious effort to find programs
that distinguish them.

However, the paper is also very different.  Where the earlier paper
focused on implementation in Scheme, Danvy and Filinski focus on
semantics.  And where Felleisen and his collaborators focused on
operational semantics, this paper provides a more denotational focus,
defining the new control operators in terms of the CPS (and ECPS)
translation for the call-by-value $\lambda$-calculus.  However, this
paper also provides a number of examples of the control operators in
action, and provides a hierarchy of such operators.

{\tt shift} and {\tt reset} have seen significantly more use in actual
programming than the other delimited continutation operators.  For
example, Filinski uses them to implement a representation of monads
\cite{monads}, and Sperber and Gasbicher provide a direct
implementation in \cite{581504}.  However, they still have not
achieved even the popularity of {\tt call/cc}.

\subsection*{Continuations and Concurrency}
Robert Hieb and R. Kent Dybvig 1990 \cite{hieb90continuations}

This paper introduced a new delimited continuation operator, {\tt
spawn}, with significant differences from those seen previously.
First, this operator was introduced with the specific purpose of
handling concurrent process control.  However, the concurrency aspects
have been neglected in the further literature.  Also, instead of being
a pair of operators, there is only one.  The operator creates a new
control operator each time it is invoked, and passes this new operator
to its argument.  Further, the newly created control operator is
delimited by the particular {\tt spawn} where it was created.  This
new operator behaves identically to {\tt control} from
\cite{sitaram90control}, with the appropriate delimiter.

In addition to defining the new operator, the authors provide a
semantics for it.  To accomplish this, they add a labeling construct
to the standard context based reduction semantics seen earlier.  This
labeling construct allows jumps back to previously labeled points in
the program execution, saving and restoring the context.  While this
operator could have been useful for defining the semantics of control
operators, it has not seen further use.

\subsection*{A Dynamic Extent Control Operator for Partial
  Continuations}
Christian Queinnec and Bernard Serpette 1991 \cite{queinnec91dynamic}

In this paper, Queinnec and Serpette propose a control operator, {\tt
splitter}, very much like {\tt spawn}.  Again, it takes the form of a
delimiter which binds a new control operator each time it is invoked.
However, in this case, two new control operators are created.  One
reifies the current continuation back to the delimiter, but does not
involve aborting the current context.  The other merely aborts the
context back to the delimiter.  These are of course the two operations
performed together in all the other control operators.  Again, lexical
scoping plays a significant role, as it did for {\tt spawn}.
Unfortunately, no semantics are provided in the paper (although an
ACPS semantics is provided in a later survey paper \cite{queinnec93library}).

This paper represents a different direction from the main stream of
control operator research.  {\tt splitter} divides the two main
effects of continuation invocation in a way not seen elsewhere.  This
makes {\tt splitter}-created continutations even more ``composable''
than those created by {\tt shift}, which are often described as
``composable''.  Further, none of the other systems involve a
specialized delimited abort construct.  (Queinnec and Moreau devised a
similar control operator in \cite{luc94partial}).

The lexically scoped control operators are fundamentally different
from dynamically scoped operators.  It is not obvious which is the
correct choice.  \footnote{Note that this distinction is irrelevant
  for {\tt call/cc}, which is not scoped.}  Lexical scope has won out
for variable binding in all modern programming languages.  However,
exceptions in languages from C++ to SML are dynamically scoped.  This
suggests that higher order control operators be scoped similarly to
their first order cousins.  Finally, it is possible to provide a
purely functional implementation of lexically scoped control
operators in terms of dynamically scoped operators.  However, the
reverse is not the case.  These reasons together suggest that dynamic
scope is the correct choice.  However, in the abscence of extensive
programming experience with such operators, such a judgement can only
be provisional.

\subsection*{Handling Control}
Dorai Sitaram 1993 \cite{sitaram93handling}

This paper introduces yet another very different control operator,
{\tt fcontrol}, athough with a confusingly similar name to some
operators that we have seen before. Here, instead of the control
operator reifying the current continuation and passing that to its
argument, as all the previous operators did, the delimiter contains a
handler procedure.  When the control operator is invoked, control
passes to the handler, which recieves both the value supplied to the
control operator and the continuation at the point the operator was
invoked (back to the delimiter).  The handler is then free to do as it
wishes with these two arguments.

This is a radical departure from control operators previously
discussed here, but it is very like an exception handler.  When an
exception is raised, control passes to the dynamically nearest
enclosing handler, which is given the value of the exception.  Here,
the addition is to provide the current continuation of the exception
invocation as well.  This allows, for example, for resumption of the
computation with a different value.

This paper further provides a number of examples (including the
classic same-fringe) using the new operator.  In these examples, we
see that much of the computation acutally takes place in the handler,
rather than in the ``body'' of the code.  This is unlike normal
programming style with either exception handlers or standard control
operators.  This paper does not provide semantics for the new
operators, however such semantics are available in
\cite{sitaram94thesis}.

\bibliography{control}

\end{document}