### 2004-09-29 classes/theory/ps1.tex

\documentclass[12pt]
{article}

\usepackage{amssymb}
\usepackage{amsmath}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{problem}[theorem]{Problem}

\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}

\title{CSG 713: Problem Set 1}
\author{Sam Tobin-Hochstadt}
\date{\today}

\begin{document}

\maketitle

\subsection*{Problem 1}

Consider the regular language $M$, accepted by a DFA $D_M =
(Q,\Sigma,\delta,q_0,F)$. Then let NFA $N_{M^R} =
(Q \cup \{q_0'\},\Sigma,\delta',q_0',\{q_0\})$, where $q_0'$ is a new
state and
$\delta' = \{ (m,n,a) ~|~ (n,m,a) \in \delta \} \cup \{
(q_0',f,\epsilon) ~|~ f \in F\}$.

To see that these two automata accept the same language, consider what
happens to each when processing a string.  For each string $w$, if there
is a path from the start state to a final state labeled with that
string in the DFA, then there must be a path from one of the states in
$F$ to $q_0$ labeled with $w^R$ in the NFA.  By construction,
therefore, there is path labeled with $\epsilon w^R$ from the start
state to a final state in the NFA.  Therefore, all strings in $W^R$
are accepted by $N_{M^R}$.

To see that all strings accepted by $N_{M^R}$ are in $W^R$, consider
what would happen if we applied the construction to a DFA with only
one start state, and added no epsilon transitions.  Then, if we
applied the construction again, we would get the original DFA.
Therefore, since the final automata accepts every string in the
reverse of the reverse of the first automata, as proved above, since
it clearly accepts no more strings than the first automata (since they
are isomorphic), the construction must not add any strings to the
language.

Finally, since NFAs can be made to have single start states without
loss of generality, and the construction also works for NFAs, the two
automata accept the same language. \qed

\subsection*{Problem 2}

\subsubsection*{Part A}

We solve this problem by proving that $C^R$ is regular.

For this problem, we need merely to keep track of whether there is a
carry bit or not, which means we need two states.  If the top is equal
to the bottom, plus a carry, then we move the the carry state.  If the
top plus the carry is equal to the bottom, we move the even state.  If
the top plus the carry is equal to the bottom plus a carry, we stay in
the carry state.  If the top is equal to the bottom, and there is no
carry, we stay in the same state.  Otherwise the machine fails.

This is implemented in an attached machine.

\subsubsection*{Part B}
Since $0 * 3^n$ and $2 * 3^n$ are even for all values of $n$, and $1 *
3^n$ is odd for all positive values of $n$, we merely need to
determine if the number of 1's in the string is even.  See the
attached figure for the automaton.

\subsection*{Problem 3}

\subsubsection*{Part A \small{I discussed this problem with Richard Cobbe}}

First, we show that multiplication by two yields a regular language.
This is trivial, since to multiply by two in binary we need merely to
shift left by one position, and compare for equality.  To do this with
an automata, we need merely to require an initial zero in the top row
when examining the string in reverse.

Since we also already are able to check for equality between the sum
of two rows and a third (Problem 2.A), we need merely to show that this combination
preserves regularity. To do this, consider a DFA that checks for
equality, and one that checks for being twice the size, when a zero is
added (this is multiplication by two).  If both accept, then accept.
This is possible because regular languages are closed under
intersection.

\subsubsection*{Part B}

For this problem, we need merely to see which row has a one where the
other has a zero first.  That row is the larger.  The automaton
implementing this is attached.

\subsection*{Problem 4}

First, note that if $(Q,\Sigma,\delta,q_0,F)$ is an NFA which accepts
language $M$, then it can also be considered as an all-paths-NFA,
accepting language $M'$.  To construct a DFA that accepts the same
language, we follow the construction for simulating an NFA with a
DFA.  However, we add one additional state, a dead state, with edges
from itself to itself on all inputs.  We add an edge to this state for
each time when the NFA would be unable to make a transition.  Further,
we restrict the accepting states to those subsets of $Q$ which are
also subsets of $F$.

This construction correctly simulates an all-paths-NFA for the same
reason that the construction works for NFAs.  If the NFA never
encounters a state where it is unable to make a transitions, then it
ends up in the state representing all possible ending states (this is
identical to the NFA case).  If all of those states are accepting, the
DFA will accept, which is the correct behavior for the
all-paths-NFA. If the all-paths-NFA ever reached a point where it was
unable to make a transition, the DFA would go to the dead state, and
reject, which is also the correct behavior.

Second, note that every DFA is an all-paths-NFA, with no
nondeterminism.  Therefore, any language that can be accepted by a DFA
can also be accepted by an all-paths-NFA.  Therefore, all-paths-NFAs
can accepts all regular languages.

From these two facts, all-paths-NFAs accepts precisely the regular
languages.  \qed

\subsection*{Problem 5 \small{I worked on this problem with Greg Pettyjohn}}
Consider a DFA $M$ which accepts language $A$.  Construct a new DFA
$M'$ by changing $F$ to $\overline{F}$.  This DFA accepts all proper
prefixes of $A$.  Call this operation $PP(A)$.

\subsubsection*{Part A}
$NOPREFIX(A) = A - PP(A)$, therefore it is regular, since the regular
languages are closed under set difference.

\subsubsection*{Part B}
$NOPREFIX(A) = \overline{PP(A)}$, therefore it is regular, since the regular
languages are closed under set complement.

\subsection*{Problem 6}

To prove that $Multiple_5$ is regular, we prove the $Multiple_5^R$ is
regular.  To do this, we construct an automaton.

Let $Q = \{0,1,2,3,4\} x \{1,2,4,3\}$.

Let $q_0 = (0,1)$.

Let $F = \{(0,x)~|~ x \in \{1,2,4,3\}\}$.

Let $\Sigma = \{0,1\}.$

Finally, we must construct the transition function.  For all states
labelled $(x,1)$, add an edge labeled 0 to the state $(x,2)$.
Similarly from $(x,2)$ to $(x,4)$ and from $(x,4)$ to $(x,3)$ and
finally from $(x,3)$ back to $(x,1)$.

Then add an edge for each state labeled $(m,n)$, add an edge labeled 1
to the state $(m+n\bmod 5, f(n))$, where $f$ takes 1 to 2, 2 to 4, 4
to 3 and 3 to 1, as in the above paragraph.

This automaton keeps track both of the number modulo 5, and how many
to add to the number if the next digit is a one.  Since the powers of
2 modulo 5 repeat the sequence $1,2,4,3$, we have a total of 20
states.  If the computation ends with the current modulus at 0, the
machine accepts.

\subsection*{Problem 7}

\subsubsection*{Part A}
The automaton is equivalent to the RE $((0 \cup 1)^*1) \cup ((0 \cup
1)^*1(00)^*)$.

\subsubsection*{Part B}
The automaton is equivalent to the RE $(b^*a) (((a \cup b)a) \cup ((a
\cup b)bb^*a))^*$.

\end{document}