### 2004-09-29 classes/theory/ps2.tex

\documentclass[12pt]
{article}

\usepackage{amssymb}
\usepackage{amsmath}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{problem}[theorem]{Problem}

% General
\def\credittype{}
\def\RC{\credittype{RC}}
\def\EC{\credittype{EC}}
\def\NA{\credittype{NA}}

% Machines
\def\machine{\sf}
\def\mdfa{{\machine DFA}}
\def\mnfa{{\machine NFA}}
\def\mgnfa{{\machine GNFA}}
\def\mcfg{{\machine CFG}}
\def\mcfl{{\machine CFL}}
\def\mpda{{\machine PDA}}

% Grammar things
\newcommand\mygrammar[1]{
$$
\begin{array}{rcll}
#1
\end{array}
$$
}
\newcommand\gr[3]{\ensuremath{{\mit #1} &\rightarrow& #2 & ~~~~;~#3} \\}
\newcommand\gb[3]{\ensuremath{&|& #2 & ~~~~;~#3} \\}
\newcommand\dr[2]{\ensuremath{{\mit #1} &\Rightarrow& #2 &} \\}
\newcommand\db[1]{\ensuremath{&\Rightarrow& #1 &} \\}
\def\por{~|~}
\def\mt{\mathtt}

% Jay's 5 rules
\newcommand\jaarule[1]{\ensuremath{\mbox{{\sc JaysRule{#1}}}}}

% # in tt
\def\h{\ensuremath{\mt{\#}}}

% Graph stuff
\newcommand\pedge[3]{\ensuremath{#1,#2 \rightarrow #3}}
\def\pedgexxx{\pedge}
\newcommand\pedgeexx[2]{\pedge{\epsilon}{#1}{#2}}
\newcommand\pedgexex[2]{\pedge{#1}{\epsilon}{#2}}
\newcommand\pedgexxe[2]{\pedge{#1}{#2}{\epsilon}}
\newcommand\pedgexee[1]{\pedge{#1}{\epsilon}{\epsilon}}
\newcommand\pedgeexe[1]{\pedge{\epsilon}{#1}{\epsilon}}
\newcommand\pedgeeex[1]{\pedge{\epsilon}{\epsilon}{#1}}
\newcommand\pedgeeee[0]{\pedge{\epsilon}{\epsilon}{\epsilon}}

% Non-terminals
\newcommand\nterm[1]{\ensuremath{\langle\mbox{{\sc #1}}\rangle}}

% non-terminals for 2.8
\def\NP{\nterm{NOUN-PHRASE}}
\def\VP{\nterm{VERB-PHRASE}}
\def\CN{\nterm{CMPLX-NOUN}}
\def\CV{\nterm{CMPLX-VERB}}
\def\AR{\nterm{ARTICLE}}
\def\N{\nterm{NOUN}}
\def\V{\nterm{VERB}}
\def\PP{\nterm{PREP-PHRASE}}
\def\P{\nterm{PREP}}
\def\SEN{\nterm{SENTENCE}}

% 2.21
\def\STMT{\nterm{STMT}}
\def\STMTLIST{\nterm{STMT-LIST}}
\def\IT{\nterm{IF-THEN}}
\def\A{\nterm{ASSIGN}}
\def\ITE{\nterm{IF-THEN-ELSE}}
\def\NIF{\nterm{NON-IF}}
\def\IF{\nterm{IF}}
\def\NEST{\nterm{NESTED-IF}}
\def\NNEST{\nterm{NON-NESTED-IF}}
\def\BEGIN{\nterm{BEGIN-END}}

%--
%Jeffrey Palm --> http://jeffpalm.com

\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}

\title{CSG 713: Problem Set 3}
\author{Sam Tobin-Hochstadt}
\date{\today}

\begin{document}

\maketitle

For all grammars, $S$ is the start symbol.  Jeff Palm supplied the
\LaTeX{} macros that made this typesetting possible.

\subsection*{Problem 1}

\subsubsection*{Part A}

\mygrammar{  
  \gr{S}{aSb  \por aSbb \por \epsilon}{a^i b^j ~with~ i \leq j \leq 2i}  
}

\subsubsection*{Part B}

\mygrammar{
  \gr{R}{aRa \por bRb \por \epsilon}{x \in \{a,b\}^*, x = x^R}
  \gr{R_2}{aR_2a \por bR_2b \por \h A\h}{x\h A^*\h x^R}
  \gr{A}{Aa \por Ab \por A\h \por \epsilon}{\{a,b,\h\}^*}
  \gr{S_1}{A\h R_2\h A \por R_2}{A^*\h R_2\h A^*}
  \gr{S}{R \por S_1}{S_1 ~or ~a ~palindrome}
}

\subsection*{Problem 2}

See PDA diagrams on attached sheet.

\subsection*{Problem 3}
Twice as many a's as b's.
 \mygrammar{
   \gr{S}{ S_1 \por S_2 \por S_3 \epsilon }{}
   \gr{S_1}{Saba \por aSba \por abSa \por abaS}{}
   \gr{S_2}{Saab \por aSab \por aaSb \por aabS}{}
   \gr{S_3}{Sbaa \por bSaa \por baSa \por baaS}{}
}

\subsection*{Problem 4}
The complement of $a^nb^n$.

\mygrammar{
  \gr{S}{X \por Y}{a^nb^n~ complement}
  \gr{X}{aXb \por A \por B}{a^nb^m, m \neq n}
  \gr{A}{Aa \por a}{a^*}
  \gr{B}{Bb \por b}{b^*}
  \gr{Y}{ZbZaZ}{a^*b^* ~complement}
  \gr{Z}{Za \por Zb \por \epsilon}{\{a,b\}^*}
}

\newpage

\subsection*{Problem 5}
Two different derivations of ``the girl touches the boy with the
flower''.

\subsubsection*{First Derivation}
Meaning: The girl uses the flower to touch the boy.

\mygrammar{  
  \dr{\SEN}{\NP \; \VP}  
  \db{\CN \; \VP}  
  \db{\AR \; \N \; \VP}  
  \db{\mt{the} \; \N \; \VP}  
  \db{\mt{the} \; \mt{girl} \; \VP}  
  \db{\mt{the} \; \mt{girl} \; \CV \; \PP}
  \db{\mt{the} \; \mt{girl} \; \V \;\NP\; \PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\NP\; \PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\AR\;\N\; \PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\N\; \PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \P\;\NP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \mt{with}\;\NP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \mt{with}\;\AR\;\N}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \mt{with}\;\mt{the}\;\N}
  \db{\mt{the} \; \mt{girl} \; \mt{touches} \;\mt{the}\;\mt{boy}\; \mt{with}\;\mt{the}\;\mt{flower}}
}

\newpage
\subsubsection*{Second Derivation}
Meaning: The girl touches the boy who has the flower.

\mygrammar{  
  \dr{\SEN}{\NP \; \VP}  
  \db{\CN \; \VP}  
  \db{\AR \; \N \; \VP}  
  \db{\mt{the} \; \N \; \VP}  
  \db{\mt{the} \; \mt{girl} \; \VP}  
  \db{\mt{the} \; \mt{girl} \; \CV}
  \db{\mt{the} \; \mt{girl} \; \V\;\NP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\NP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\CN\;\PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\AR\;\N\;\PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\N\;\PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\PP}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\P\;\CN}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\mt{with}\;\CN}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\mt{with}\;\AR\;\N}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\mt{with}\;\mt{the}\;\N}
  \db{\mt{the} \; \mt{girl} \; \mt{touches}\;\mt{the}\;\mt{boy}\;\mt{with}\;\mt{the}\;\mt{flower}}

}

\subsection*{Problem 6}

The following string has two derivations:

$\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \mt{a:=1}$

\subsubsection*{First derivation}

\mygrammar{
\dr{\STMT}{\IT}
\db{\mt{if ~condition~ then} \; \STMT}
\db{\mt{if ~condition~ then} \; \ITE}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \STMT \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \A \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \A}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \mt{a:=1}}
}

\subsubsection*{Second derivation}

\mygrammar{
\dr{\STMT}{\ITE}
\db{\mt{if ~condition~ then} \; \STMT \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \IT\; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \STMT \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \A \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \STMT}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \A}
\db{\mt{if ~condition~ then} \; \mt{if ~condition~ then} \; \mt{a:=1} \; \mt{else}\; \mt{a:=1}}
}

\subsection*{Problem 7}

An unabiguous grammar for statements.

\mygrammar{
\gr{S}{ITE \por S'}{}
\gr{S'}{IT \por A \por BE}{}
\gr{A}{\mt{a:=1}}{}
\gr{BE}{\mt{begin~} \; SL \; \mt{~end}}{}
\gr{SL}{SL \; S \por S}{}
\gr{IT}{\mt{if ~condition~} \; S \mt{~then~} S'}{}
\gr{IT}{\mt{if ~condition~} \; S \mt{~then~} S \mt{~else~} S}{}
}

\subsection*{Problem 8}

\mygrammar{
  \gr{S}{0S0 \por 1S1 \por 0S'1 \por 0S'0}{x\h y, |x| = |y|, x \neq y}
  \gr{S'}{AS'A \por \h}{x\h y, |x| = |y|}
  \gr{A}{1 \por 0}{1 ~or~ 0}

}

\subsection*{Problem 9}

\mygrammar{
  \gr{S}{0S0 \por 1S1 \por 0S'1 \por 0S'0}{xy, |x| = |y|, x \neq y}
  \gr{S'}{AS'A \por \epsilon}{xy, |x| = |y|}
  \gr{A}{1 \por 0}{1 ~or~ 0}

}

\end{document}