### 2004-12-12 classes/algorithms/final.tex
work on final exam

\documentclass[twoside]{amsart}
\usepackage{geometry}
\usepackage{stmaryrd}
\usepackage{cancel}
\usepackage{clrscode}
\usepackage{textcomp}
\geometry{tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage[curve]{xypic}

\newsavebox{\fmbox}
\newenvironment{fmpage}[1]
{\begin{lrbox}{\fmbox}
   \begin{minipage}{#1}}
{\end{minipage}
\end{lrbox}\fbox{\usebox{\fmbox}}}

\newenvironment{directive}{\begin{fmpage}{5.5in}}{\end{fmpage}}
\newenvironment{problem}[1]{\newpage\section{#1}\begin{fmpage}{5.5in}}{\end{fmpage}}

\newtheorem{prop}{Proposition}
\newtheorem{claim}[prop]{Claim}
\newtheorem{lemma}[prop]{Lemma}

\begin{document}

{\raggedleft
Sam Tobin-Hochstadt \\
CSG 713 Final Exam \\
\today \\
{\tt samth@ccs.neu.edu} \\
}

%\newcommand{\min}[#1]{\textrm{min}(

\section*{Problem 1}

First, initialize min($v_1$) to $v_1$.  Then, for each successive
vertex $v_k$, perform BFS from that vertex.  If a previously finished
vertex $v_j$ is reached, then terminate the BFS and min($v_k$) =
min($v_j$). This algorithm visits every vertex and every edge at most
once, and thus takes $O(V + E)$ time.

\section*{Problem 2}

\subsection*{Part i}
This conjecture is false.

\xymatrix{
  v_1 \ar@{-}[d] \ar@{-}[r] \ar@{-}[dr] & v_2 \ar@{-}[d] \\
  v_3 \ar@{-}[r] & v_4
}

\subsection*{Part ii}
This conjecture is false.

\xymatrix{
  v_1 \ar@{-}[rr] \ar@{-}[ddr] \ar@{-}[dr]  & \txt{} & v_2 \ar@{-}[ddl]
  \ar@{-}[dl]    \\
  \txt{} & v_3 \ar@{-}[d]   & \txt{} \\
  \txt{} & v_4    & \txt{} 
}

\section*{Problem 3}

\subsection*{Part i}

This conjecture is true.

\subsection*{Part ii}

This conjecture is false.

\xymatrix{
  v1 \ar@{-}[r] \ar@{-}[d] & v2 \\
  v3 \ar@{-}[r] & v4 \ar@{-}[u]
}

\section*{Problem 4}

The Laksurk method successfully finds a minimum spanning tree for any
graph.

\section*{Problem 5}

\section*{Problem 6}

\subsection*{Part i}

\subsection*{Part ii}

Create a vertex for every number in the sequence.  Then create an edge
from every vertex to

\section*{Problem 7}

\section*{Problem 8}

To see that an approximation ratio of 2 is tight, consider the
following graph:

\xymatrix{
  v_1 \ar@{-}[r] & v_3 \ar@{-}[r] & v_2
}

Then $A$ and $B$ will be initialized, and there will only be one
remaining vertex.  Since $d(v_3,\{v_1\}) = d(v_3,\{v_2\}) = 1$, $v_3$
will be allocated to $B$.  This produces a cut of size 1.  However,
the maximal cut is $(\{v_1,v_2\},\{v_3\})$, which has size 2.

\end{document}